\section{Reproducibility}
\label{app:repro}

This appendix describes how to rebuild the paper and reproduce the ordering evidence. The goal is traceability and comparable regimes, not identical milliseconds on different hardware.

\subsection{Paper build}

Build from \texttt{paper/}:

\begin{verbatim}
pdflatex -interaction=nonstopmode main.tex
bibtex main
pdflatex -interaction=nonstopmode main.tex
pdflatex -interaction=nonstopmode main.tex
\end{verbatim}

The public source bundle should include \texttt{main.tex}, \texttt{template.tex}, \texttt{PRIMEarxiv.sty}, \texttt{sections/}, \texttt{figures/}, \texttt{library.bib}, and \texttt{00README.json}. If a venue template replaces the arXiv style, rebuild with that venue's class and rules.

\subsection{Repository and revision}

The implementation is open source at \url{https://github.com/botirk38/tensora}. Record the exact revision next to exported results:

\begin{verbatim}
git rev-parse HEAD
git status -sb
\end{verbatim}

Also record \texttt{date -u}, \texttt{uname -a}, \texttt{rustc --version}, \texttt{python3 --version}, GPU and driver information from \texttt{nvidia-smi}, and the active vLLM package version when vLLM is used.

\subsection{Rust profiling}

Build the profiler in release mode:

\begin{verbatim}
cargo build --release --bin profile
\end{verbatim}

Run one cell with:

\begin{verbatim}
cargo run --release --bin profile -- <format> <backend> \
  --model-id <MODEL_ID> --iterations 1
\end{verbatim}

\noindent where \texttt{<format>} is \texttt{safetensors} or \texttt{serverlessllm}, and \texttt{<backend>} is \texttt{sync}, \texttt{async}, or \texttt{io-uring}. SafeTensors uses the Hugging Face Hub cache. ServerlessLLM-style conversions are stored under the Tensora cache directory, for example \texttt{\$XDG\_CACHE\_HOME/tensora/<slug>/serverlessllm}.

For cold-cache runs on hosts with sufficient privileges, run \texttt{sync} followed by \texttt{echo 3 > /proc/sys/vm/drop\_caches} before each invocation. On restricted containers, use the repository-supported eviction path and report that deviation. The full H100 matrix in \path{results/h100/rust.tsv} was generated by sweeping model ids, formats, and backends with that ritual.

\subsection{Python and vLLM harnesses}

Python benchmarks live under \path{bindings/python/benchmarks/}. The convenience script \path{scripts/run\_benchmarks.sh} runs pytest-benchmark suites and writes JSON outputs. Set \texttt{TENSORA\_BENCH\_MODELS} to one or more Hugging Face model ids, separated by spaces or commas. Set \texttt{TENSORA\_BENCH\_NO\_VLLM=1} when a GPU-backed vLLM environment is unavailable.

When vLLM is enabled, runs are serialized by model to avoid GPU memory interference. The paper's archived exports live under \path{results/h100/}, \path{results/h200/}, and \path{results/a100/}. The H100 vLLM matrix in \path{results/h100/vllm.tsv} sweeps loaders, cache modes, and benchmark kinds: \texttt{load\_only}, \texttt{TTFT}, and \texttt{steady\_state\_decode}.

\subsection{What a replication should check}

A strong replication should check ordering regimes before exact milliseconds:
\begin{itemize}
\item SafeTensors changes backend preference with scale.
\item ServerlessLLM-style partitioning does not follow the SafeTensors sync-first rule.
\item vLLM initialization is sensitive to loader choice and includes Python integration overhead from the binding layer, while steady-state decode is much less sensitive.
\item Blocking \texttt{io\_uring} changes large-scale winners.
\end{itemize}

\section{Cold-cache anchor statistics}
\label{app:anchor-stats}

Tables~\ref{tab:anchor-safetensors} and \ref{tab:anchor-serverlessllm} summarize five cold-cache repetitions for all 18 anchored H100 cells from \path{results/h100/anchor.tsv}. Anchors cover Qwen3-0.6B, Qwen3-8B, and Qwen3-32B across all three backends and both formats (18 of 36 total matrix cells). The main result tables use one representative cold run per cell for readability; these anchors provide statistical support.

The \emph{Sep.}\ column indicates whether the backend's range is cleanly separated from the next-closest backend at the same model/format ($\checkmark$ = non-overlapping, $\circ$ = overlapping with at least one comparator). Where ranges overlap, the ranking should not be treated as strict.

\begin{table}[htbp]
\centering
\small
\caption{SafeTensors anchor repetitions: median (ms) with range across five cold-cache runs.}
\label{tab:anchor-safetensors}
\begin{tabular}{llrrrl}
\toprule
Model & Backend & Median & Range & CV\% & Sep. \\
\midrule
Qwen3-0.6B & sync & 323 & 322 to 403 & 9.1 & $\checkmark$ \\
Qwen3-0.6B & io\_uring & 1\,159 & 1\,156 to 1\,264 & 3.6 & $\checkmark$ \\
Qwen3-0.6B & async & 1\,735 & 1\,572 to 1\,792 & 4.6 & $\checkmark$ \\
\midrule
Qwen3-8B & io\_uring & 3\,250 & 3\,160 to 3\,274 & 1.3 & $\checkmark$ \\
Qwen3-8B & sync & 3\,701 & 3\,648 to 3\,731 & 0.8 & $\circ$ \\
Qwen3-8B & async & 6\,805 & 6\,510 to 7\,161 & 3.3 & $\checkmark$ \\
\midrule
Qwen3-32B & io\_uring & 12\,204 & 11\,737 to 14\,389 & 7.6 & $\checkmark$ \\
Qwen3-32B & sync & 15\,462 & 15\,327 to 16\,144 & 2.0 & $\circ$ \\
Qwen3-32B & async & 26\,685 & 25\,868 to 27\,416 & 1.9 & $\checkmark$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[htbp]
\centering
\small
\caption{ServerlessLLM anchor repetitions: median (ms) with range across five cold-cache runs.}
\label{tab:anchor-serverlessllm}
\begin{tabular}{llrrrl}
\toprule
Model & Backend & Median & Range & CV\% & Sep. \\
\midrule
Qwen3-0.6B & async & 1\,353 & 1\,331 to 1\,409 & 2.1 & $\circ$ \\
Qwen3-0.6B & io\_uring & 1\,417 & 1\,395 to 1\,483 & 2.2 & $\circ$ \\
Qwen3-0.6B & sync & 1\,456 & 1\,409 to 1\,524 & 3.0 & $\circ$ \\
\midrule
Qwen3-8B & io\_uring & 7\,267 & 7\,119 to 8\,885 & 8.8 & $\circ$ \\
Qwen3-8B & async & 9\,172 & 7\,628 to 9\,837 & 10.0 & $\circ$ \\
Qwen3-8B & sync & 10\,525 & 10\,175 to 10\,676 & 1.9 & $\checkmark$ \\
\midrule
Qwen3-32B & io\_uring & 27\,910 & 26\,664 to 29\,152 & 2.9 & $\circ$ \\
Qwen3-32B & async & 29\,011 & 28\,639 to 29\,222 & 0.8 & $\circ$ \\
Qwen3-32B & sync & 39\,241 & 37\,785 to 40\,691 & 2.4 & $\checkmark$ \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Key observations from anchor analysis.} The anchors confirm the paper's main structural claims with non-overlapping ranges: (1)~SafeTensors io\_uring dominates sync at 8B and 32B, (2)~sync dominates io\_uring at 0.6B, and (3)~sync is the slowest ServerlessLLM backend at all anchored scales. Within ServerlessLLM, the ordering among async and io\_uring is not robustly separable at any individual scale; the consistent finding is that both outperform sync, not that either one dominates.

\section{Modal anchor statistics}
\label{sec:modal-anchors}

A five-repetition matrix was collected on Modal H100 containers (gVisor, \texttt{io\_uring} blocked). This matrix covers all six models across sync and async backends, with close-call analysis derived from this matrix. Table~\ref{tab:modal-safetensors-anchors} and Table~\ref{tab:modal-serverlessllm-anchors} summarize the SafeTensors and ServerlessLLM results.

\begin{table}[htbp]
\centering
\small
\caption{Modal H100 SafeTensors anchor repetitions: median (ms) with range across five cold-cache runs.}
\label{tab:modal-safetensors-anchors}
\begin{tabular}{llrrrr}
\toprule
Model & Backend & Median & Range & CV\% & $N$ \\
\midrule
Qwen3-0.6B & async & 558 & 518--635 & 8.2 & 5 \\
Qwen3-0.6B & sync & 713 & 675--724 & 3.0 & 5 \\
\midrule
SmolLM3-3B & async & 2\,339 & 2\,120--2\,547 & 6.6 & 5 \\
SmolLM3-3B & sync & 2\,602 & 1\,647--2\,832 & 19.5 & 5 \\
\midrule
Qwen3-4B & async & 2\,908 & 2\,764--3\,131 & 5.4 & 5 \\
Qwen3-4B & sync & 3\,182 & 1\,813--3\,232 & 21.2 & 5 \\
\midrule
Qwen3-8B & async & 5\,378 & 3\,576--6\,267 & 22.2 & 5 \\
Qwen3-8B & sync & 6\,496 & 3\,781--6\,502 & 20.2 & 5 \\
\midrule
Qwen3-14B & async & 9\,910 & 6\,681--10\,748 & 21.7 & 5 \\
Qwen3-14B & sync & 10\,913 & 6\,647--11\,189 & 19.2 & 5 \\
\midrule
Qwen3-32B & async & 20\,871 & 14\,418--23\,550 & 21.7 & 5 \\
Qwen3-32B & sync & 23\,314 & 15\,191--24\,039 & 17.1 & 5 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[htbp]
\centering
\small
\caption{Modal H100 ServerlessLLM anchor repetitions: median (ms) with range across five cold-cache runs.}
\label{tab:modal-serverlessllm-anchors}
\begin{tabular}{llrrrr}
\toprule
Model & Backend & Median & Range & CV\% & $N$ \\
\midrule
Qwen3-0.6B & async & 1\,171 & 1\,118--1\,206 & 3.3 & 5 \\
Qwen3-0.6B & sync & 1\,303 & 1\,068--1\,402 & 9.9 & 5 \\
\midrule
SmolLM3-3B & async & 4\,569 & 4\,263--4\,684 & 3.5 & 5 \\
SmolLM3-3B & sync & 4\,741 & 4\,272--5\,221 & 7.1 & 5 \\
\midrule
Qwen3-4B & async & 6\,233 & 5\,494--7\,058 & 9.4 & 5 \\
Qwen3-4B & sync & 6\,563 & 5\,138--7\,079 & 11.9 & 5 \\
\midrule
Qwen3-8B & async & 12\,808 & 10\,835--13\,579 & 8.4 & 5 \\
Qwen3-8B & sync & 13\,382 & 11\,369--13\,872 & 8.4 & 5 \\
\midrule
Qwen3-14B & sync & 22\,816 & 21\,697--24\,871 & 5.4 & 5 \\
Qwen3-14B & async & 23\,069 & 22\,043--23\,327 & 2.4 & 5 \\
\midrule
Qwen3-32B & async & 51\,720 & 47\,340--56\,205 & 6.1 & 5 \\
Qwen3-32B & sync & 53\,971 & 28\,707--71\,612 & 38.4 & 5 \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Key observations.} (1)~Sync is the slowest or near-slowest ServerlessLLM backend at all Modal scales, replicating the primary H100 finding. (2)~SafeTensors at 0.6B shows a clean separation: async (558\,ms) $<$ sync (713\,ms); without \texttt{io\_uring}, async becomes the small-scale winner rather than sync. (3)~Most mid-to-large SafeTensors and ServerlessLLM cells show overlapping ranges among non-sync backends, consistent with the H100 observation that close-call rankings are noise-dominated. (4)~CDN-induced variance (Qwen3-32B sync CV=38\%) inflates ranges on Modal relative to local-NVMe H100; regime structure is preserved, but individual-cell precision is lower.

\textbf{Close-call analysis.} The full anchor matrix confirms range overlap at the mid-to-large cells where backend differences narrow. SafeTensors Qwen3-14B: sync [6\,647 to 11\,189\,ms] vs.\ async [6\,681 to 10\,748\,ms] (\textbf{OVERLAP}). ServerlessLLM Qwen3-14B: sync [21\,697 to 24\,871\,ms] vs.\ async [22\,043 to 23\,327\,ms] (\textbf{OVERLAP}). ServerlessLLM Qwen3-4B: sync [5\,137 to 7\,079\,ms] vs.\ async [5\,494 to 7\,058\,ms] (\textbf{OVERLAP}). These confirm that the differences at those cells are not statistically separable, consistent with the paper's conservative separation criterion.


\section{Cross-environment validation}
\label{app:cross-env}

This section presents the full cross-environment matrices referenced in Section~\ref{sec:cross-env}. Results are collected on H200 and A100 container environments where \texttt{io\_uring} is blocked at the kernel level (\texttt{EPERM}).

\subsection{H200 Results (\texttt{io\_uring} blocked)}

Table~\ref{tab:h200-safetensors} shows the H200 SafeTensors matrix. With \texttt{io\_uring} blocked, async replaces it as the fastest backend at large scale, while sync remains dominant for small models.

\begin{table}[htbp]
\caption{H200 cold-cache SafeTensors results (ms). \texttt{io\_uring} is blocked (\texttt{EPERM}). Sync dominates at small scale; async takes over at large scale.}
\label{tab:h200-safetensors}
\centering
\begin{tabular}{lrr}
\toprule
Model & Sync & Async \\
\midrule
Qwen3-0.6B & 604.34 & 852.17 \\
SmolLM3-3B & 2013.66 & 2938.30 \\
Qwen3-4B & 909.83 & 2478.10 \\
Qwen3-8B & 3924.69 & 2977.59 \\
Qwen3-14B & 25872.03 & 4767.67 \\
Qwen3-32B & 36205.91 & 10623.83 \\
\bottomrule
\end{tabular}
\end{table}

The H200 SafeTensors crossover preserves the structural pattern from H100: sync wins at small scale (0.6B to 4B), while a different backend dominates at large scale (8B+). On H100 that backend is \texttt{io\_uring}; on H200 it is async.

Table~\ref{tab:h200-serverlessllm} shows the H200 ServerlessLLM matrix.

\begin{table}[htbp]
\caption{H200 cold-cache ServerlessLLM results (ms). Async dominates all rows.}
\label{tab:h200-serverlessllm}
\centering
\begin{tabular}{lrr}
\toprule
Model & Sync & Async \\
\midrule
Qwen3-0.6B & 1565.27 & 1491.44 \\
SmolLM3-3B & 3945.68 & 1947.39 \\
Qwen3-4B & 5100.25 & 2489.27 \\
Qwen3-8B & 10788.88 & 6045.47 \\
Qwen3-14B & 17765.28 & 14450.40 \\
Qwen3-32B & 39061.57 & 27367.74 \\
\bottomrule
\end{tabular}
\end{table}

The ServerlessLLM ordering on H200 matches H100 structurally: sync is never the fastest, and async dominates all rows.

\subsection{A100 Results (\texttt{io\_uring} blocked, reduced ladder)}

The A100 environment runs a four-model reduced ladder (0.6B, 4B, 8B, 32B) with the same \texttt{EPERM} constraint. Table~\ref{tab:a100-safetensors} and Table~\ref{tab:a100-serverlessllm} summarise the results.

\begin{table}[htbp]
\caption{A100 cold-cache SafeTensors results (ms). Network-attached storage with AMD EPYC CPU. Sync wins at 0.6B; async dominates at 4B to 8B; sync recovers at 32B.}
\label{tab:a100-safetensors}
\centering
\begin{tabular}{lrr}
\toprule
Model & Sync & Async \\
\midrule
Qwen3-0.6B & 724.26 & 1094.43 \\
Qwen3-4B & 7598.41 & 3532.11 \\
Qwen3-8B & 8412.36 & 4235.21 \\
Qwen3-32B & 46923.42 & 60807.34 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[htbp]
\caption{A100 cold-cache ServerlessLLM results (ms). Sync and async are essentially tied at 0.6B; async dominates 4B and above.}
\label{tab:a100-serverlessllm}
\centering
\begin{tabular}{lrr}
\toprule
Model & Sync & Async \\
\midrule
Qwen3-0.6B & 2183.33 & 2250.91 \\
Qwen3-4B & 7551.79 & 4486.75 \\
Qwen3-8B & 14707.01 & 11528.26 \\
Qwen3-32B & 56025.98 & 40472.54 \\
\bottomrule
\end{tabular}
\end{table}

The A100 SafeTensors results show a distinctive pattern: sync wins at 0.6B, async dominates the mid-range (4B to 8B), but sync recovers at 32B. This contrasts with both H100 (where \texttt{io\_uring} dominates large scale) and H200 (where async dominates large scale). For ServerlessLLM, sync and async are essentially tied at 0.6B (2183~ms vs 2251~ms); async dominates from 4B upward.

\subsection{Cross-Environment Summary}

Table~\ref{tab:cross-env-best} summarises the best explicit backend per row across all four environments.

\begin{table}[htbp]
\caption{Best explicit backend by environment for SafeTensors. The crossover from sync to another backend persists across all four environments. Which alternative wins at mid-to-large scale depends on hardware, storage topology, and kernel interface availability.}
\label{tab:cross-env-best}
\centering
\begin{tabular}{lllll}
\toprule
Model & H100 & H200 & A100 & Modal H100 \\
\midrule
Qwen3-0.6B & sync & sync & sync & async \\
Qwen3-4B & sync & sync & async & async \\
Qwen3-8B & io\_uring & async & async & async \\
Qwen3-32B & io\_uring & async & sync & async \\
\bottomrule
\end{tabular}
\end{table}

Four findings from the cross-environment comparison: (1)~Sync dominates small SafeTensors models in three of four environments; on Modal (gVisor), async wins even at 0.6B because \texttt{io\_uring} is blocked and the gVisor syscall path penalizes synchronous reads. (2)~At large scale, the strongest backend depends on available kernel interfaces and storage topology: \texttt{io\_uring} on local NVMe (H100), async on network storage (H200 and Modal), and sync on a different network-storage configuration (A100 at 32B). (3)~Small-scale ordering transfers better than large-scale ordering: the small-scale winner is consistent across H100, H200, and A100, while the large-scale winner varies. (4)~Modal H100 adds a fourth datapoint: the regime structure (non-sync backends dominate at mid-to-large scale) persists under gVisor sandboxing with ephemeral storage and CDN-based model fetching.

\section{Held-out model validation}
\label{app:held-out}

The primary model ladder is drawn from a single checkpoint family (Qwen3, supplemented by SmolLM3-3B). To test whether the regime structure generalises beyond this geometry, two held-out models not used in threshold derivation were evaluated on Modal H100 containers (\texttt{io\_uring} unavailable).

\begin{table}[htbp]
\caption{Held-out model validation on Modal H100 (ms, \texttt{io\_uring} blocked). Mistral-7B is a different architecture family with 3~shards and 582~tensors.}
\label{tab:held-out}
\centering
\begin{tabular}{llrr}
\toprule
Model & Format & Sync & Async \\
\midrule
Qwen3-8B & SafeTensors & 11\,330 & 3\,884 \\
Qwen3-8B & ServerlessLLM & 10\,662 & 5\,184 \\
Mistral-7B & SafeTensors & 6\,889 & 8\,596 \\
Mistral-7B & ServerlessLLM & 24\,195 & 21\,888 \\
\bottomrule
\end{tabular}
\end{table}

Two regime findings transfer. First, \textbf{sync is never the fastest ServerlessLLM backend} for either held-out model: async beats sync by $2.06\times$ on Qwen3-8B and by $1.11\times$ on Mistral-7B. The structural claim that partitioned layouts disfavor synchronous submission holds across both checkpoint families. Second, \textbf{SafeTensors at 8B scale favours non-sync backends} when \texttt{io\_uring} is unavailable: async is $2.92\times$ faster than sync for Qwen3-8B SafeTensors, consistent with the H200 finding that async replaces \texttt{io\_uring} as the large-scale winner in restricted environments. Mistral-7B SafeTensors shows a different pattern: sync is fastest ($6\,889$\,ms vs.\ async $8\,596$\,ms), consistent with its 3-shard geometry reducing the concurrency benefit that async exploits on the 5-shard Qwen3-8B.

Mistral-7B's ServerlessLLM spread is narrower ($1.11\times$) than Qwen3-8B's ($2.06\times$). This is consistent with Mistral's different shard geometry (3~shards of $\sim$9.7\,GB each vs.\ Qwen3-8B's 5~shards of $\sim$3.3\,GB), which reduces the concurrency advantage that async exploits. The regime structure transfers; the effect size is geometry-dependent.



\section{Artifact layout}
\label{app:artifact-layout}

Each results directory contains tab-separated summaries and raw exports. \path{results/<env>/rust.tsv} contains Rust cold-cache profiles. \path{results/<env>/vllm.tsv} contains vLLM integration benchmarks where available. \path{results/<env>/anchor.tsv} contains repeated anchors where collected. \path{results/<env>/raw/} contains pytest-benchmark JSON exports.

Archived exports may include exploratory or fallback rows (e.g., rows with \texttt{default} backend label from earlier versions of the study). These are retained for provenance but are not referenced in the paper's main comparison. The paper reports only the three explicit backends (sync, async, \texttt{io\_uring}) as described in Section~\ref{sec:experimental_setup}.
