\section{Conclusion}
\label{sec:conclusion}

The central finding of this paper is that checkpoint loading for LLM serving exhibits identifiable performance regimes: the optimal I/O backend changes systematically with checkpoint format, model scale, and deployment environment. This is not a single bandwidth number. It is a workload-shaped systems problem. The layout determines request geometry, the backend determines submission policy, runtime capability determines which interfaces can be used, and the integration layer determines how much of a backend improvement reaches users.

Tensora measures this path across two layouts, four backend policies, three observation layers, six public checkpoints, and four environments. The results reject a universal backend. H100 SafeTensors shifts from sync-favored small loads to \texttt{io\_uring}-favored large loads. H100 ServerlessLLM-style partitioning eliminates sync as a winner and rewards grouped range execution and planning. Python and vLLM reorder the details but preserve the practical consequence: loader choice changes public API time and serving initialization. H200, A100, and Modal H100 show that regime structure transfers better than backend identity. Held-out validation on Mistral-7B-Instruct---a different architecture family with different shard geometry---confirms that the structural claims (sync-never-wins for ServerlessLLM, non-sync backends dominate SafeTensors at 8B+) hold beyond the primary model ladder, with geometry-dependent effect sizes. An external baseline (\texttt{hf-native}: upstream SafeTensors crate) contextualises Tensora's contribution: on environments with \texttt{io\_uring}, the best Tensora backend outperforms the upstream reference by $1.4$--$1.5\times$ at mid-to-large scale.

The recommendation is therefore simple and bounded: load by design. Select from measured workload features, gate the choice by runtime capability, and expose explicit overrides for boundary cases. The right abstraction is not an I/O API name; it is a format-aware adaptive policy that recognizes checkpoint loading as a workload-dependent decision with measurable regime boundaries.
