\section{Discussion}
\label{sec:discussion}

The main result is not that one backend wins. The main result is that a checkpoint loader must treat layout and scale as policy inputs.

\subsection{Mechanism: why layout changes the answer}

SafeTensors and ServerlessLLM-style partitioning create different I/O problems. SafeTensors exposes large contiguous reads after metadata parsing. For small checkpoints, a practical sync backend has low overhead and enough parallelism; additional scheduling machinery does not amortize. As size and shard count grow, the H100 workload has enough in-flight work for \texttt{io\_uring} to pay off.

ServerlessLLM-style partitioning changes the request geometry. The loader reconstructs tensors from indexed partition files, so grouping, coalescing, and range batching dominate. The SafeTensors rule ``sync can be good for small loads'' does not transfer: sync never wins in the H100 ServerlessLLM matrix. The format, not the API name, explains the reversal.

\subsection{Integration: why rankings move across layers}

Rust, Python, and vLLM answer different questions. Rust isolates backend mechanics. Python measures a public library boundary with FFI, tensor conversion, and object construction. vLLM measures serving startup with CUDA context setup, memory pools, runtime configuration, and generation setup. These costs can dilute a pure I/O advantage or reorder two close backends.

This is not a flaw in the methodology; it is the deployment reality. A systems paper about loading should show both mechanism and propagation. Tensora's gains matter because they remain visible in vLLM cold initialization even after fixed engine costs are included.

\subsection{Policy: what \texttt{default} should promise}

An adaptive default should promise three bounded properties. First, it should avoid clearly wrong global choices: sync for all partitioned loads, or \texttt{io\_uring} for every tiny contiguous load. Second, it should degrade gracefully when a backend is unavailable, as in containers that block \texttt{io\_uring}. Third, it should expose explicit overrides for known boundary cases.

It should not promise oracle optimality. Table~\ref{tab:oracle} shows misses where the current feature set is too coarse. Those misses are useful because they identify where the next policy should add signals: finer shard geometry, storage topology, or learned per-environment calibration. The correct engineering response is transparency plus overrides, not hiding the selector behind a single API name.

\subsection{What transfers}

Absolute milliseconds are machine-specific. Ordering regimes transfer more plausibly. In this study, small SafeTensors loads favor sync across H100, H200, and A100. The large-scale winner does not transfer by name: H100 favors \texttt{io\_uring}, H200 favors async when \texttt{io\_uring} is blocked, and A100 shows a non-monotonic recovery at the largest SafeTensors row. A replication should therefore ask whether workload-shape changes persist, not whether every number matches.

\subsection{A100 SafeTensors at 32B: why sync recovers}
\label{sec:a100-anomaly}

The A100 SafeTensors result at 32B presents a distinctive reversal: sync (46,923~ms) outperforms async (60,807~ms) by $1.30\times$, while the adaptive default (43,817~ms) is the fastest path. This contrasts sharply with H200 at the same scale, where async dominates sync by $3.41\times$.

A plausible explanation involves the interaction between storage-link saturation and async scheduling overhead. The A100 environment uses AMD EPYC 7543 processors with network-attached storage whose bandwidth and latency characteristics differ from the H200's Intel Xeon 8568Y+ with its own network-attached storage. At 65~GB, the checkpoint transfer may saturate the available storage bandwidth on the A100 network path. When the storage link is the bottleneck, async's Tokio task scheduling---per-file grouping, dynamic batching, and runtime coordination---adds overhead without a corresponding concurrency benefit, because the link cannot serve additional parallel requests faster than it can serve sequential ones. Sync's simpler thread-parallel POSIX reads saturate the same link with lower scheduling cost.

This explanation is consistent with three observations. First, async outperforms sync at smaller A100 SafeTensors scales (4B, 8B), where the transfer size does not saturate the link and concurrency still yields benefit. Second, the reversal does not appear on H200, where the storage path apparently sustains higher effective bandwidth at 32B scale. Third, the adaptive default outperforms both explicit backends at A100 32B because its read planning (chunking, coalescing) optimizes transfer efficiency rather than relying solely on concurrency.

This finding illustrates why storage topology should be considered an additional policy input beyond format and model scale. On environments where network-attached storage has a hard bandwidth ceiling, the overhead of async coordination can exceed its parallelism benefit at very large transfer sizes. If the A100 environment were still available, targeted instrumentation---measuring per-backend bandwidth utilization and request latency distribution---could confirm or refine this hypothesis. Even without that instrumentation, the result motivates a useful design principle: an adaptive selector should not assume that async is unconditionally superior to sync at large scale; the answer depends on whether the storage path can actually serve concurrent requests faster.

\subsection{Security context for capability-gated backend selection}
\label{sec:security-discussion}

The capability-gated fallback described in Section~\ref{sec:design} is not merely a deployment convenience. It reflects a deliberate security posture in production container environments. The \texttt{io\_uring} subsystem has been a recurring source of kernel vulnerabilities, including privilege escalation and use-after-free bugs~\cite{iouring_cve_2022_29582,iouring_cve_2023_2598}. In response, Google's production container infrastructure restricts \texttt{io\_uring} system calls via seccomp filters~\cite{google_iouring_restriction}. Docker's default seccomp profile similarly blocks \texttt{io\_uring\_setup} and \texttt{io\_uring\_enter}~\cite{docker_seccomp_default}. The Linux kernel itself added the \texttt{io\_uring\_disabled} sysctl in version 6.6 to allow system administrators to restrict the interface globally~\cite{lwn_iouring_sysctl}.

This security context reframes the \texttt{EPERM} failures observed on H200 and A100 (Section~\ref{sec:cross-env}) as expected behavior under a security-hardened deployment, not as an accidental limitation. The capability-gated selector handles this correctly: it probes for \texttt{io\_uring} availability at startup, falls back to async or sync when the interface is blocked, and never attempts to circumvent the restriction.

For deployment-sensitive environments, the security posture of the host is therefore an additional factor influencing backend availability, alongside format and scale. A checkpoint loader that hard-codes \texttt{io\_uring} as its primary path would either fail outright in restricted containers or require the operator to weaken the container's security profile. The capability-gated approach avoids this trade-off: it uses \texttt{io\_uring} when available and safe, and selects the next-best backend otherwise.

\subsection{Checkpoint format and model supply chain security}
\label{sec:format-security}

SafeTensors was designed specifically to prevent arbitrary code execution during model loading~\cite{safetensors_security}. Traditional checkpoint formats based on Python's \texttt{pickle} protocol allow arbitrary code to run during deserialization, creating a well-documented attack vector in the model supply chain~\cite{pytorch_pickle_warning}. SafeTensors eliminates this class of vulnerability by restricting the file format to a flat header plus contiguous tensor data, with no executable content.

This security property is orthogonal to the performance considerations evaluated in this paper, but it provides additional context for the choice of SafeTensors as a primary evaluation format. In deployment environments where models are loaded from shared registries or untrusted sources, the safe deserialization guarantee is a meaningful benefit beyond loading speed. The combination of a security-safe checkpoint format with a capability-gated backend selector addresses two distinct deployment concerns: avoiding code execution during deserialization (format level) and avoiding kernel attack surface during I/O submission (backend level).

\subsection{Implications}

For practitioners, log the features that determine loader choice: format, bytes, shard or partition count, backend availability, cache condition, security policy, and integration layer. Expose a default, but also expose explicit backend labels and override hooks.

For researchers, report loading results as layered constructs. A Rust microbenchmark, a Python API benchmark, and a serving-engine startup benchmark are all valuable; none subsumes the others. A paper that reports only one layer risks explaining the wrong phenomenon.
