\section{Introduction}
\label{sec:introduction}

An LLM serving request cannot generate a token until model weights are resident. In serverless and multi-tenant settings, idle models are commonly evicted to recover scarce accelerator memory; the next request then pays a cold load before inference begins. A 7B float16 checkpoint is already on the order of 14~GB, and the checkpoints in this paper reach beyond 65~GB of weight bytes. Cold loading is therefore not a startup footnote. It is part of the user-visible serving path.

The usual shorthand is to call this a bandwidth problem. That shorthand is incomplete. A loader does not read an abstract number of gigabytes; it executes a concrete read plan. The checkpoint format decides whether bytes appear as a few contiguous shards or as many indexed ranges across partition files. The backend decides whether those reads are submitted through blocking POSIX calls, an async runtime, or \texttt{io\_uring}. The integration layer then adds its own costs: Python object construction, tensor materialization, CUDA context setup, allocator initialization, and serving-engine configuration. A device can have high peak throughput and still produce poor cold-start time when request geometry and submission policy are mismatched.

Existing LLM serving systems and checkpoint libraries typically treat loading as a fixed initialization cost with a single code path. No prior work, to our knowledge, systematically characterizes checkpoint loading as a \emph{policy selection problem} with identifiable performance regimes that shift with format, scale, and deployment environment. This gap matters because the optimal backend is not stable: it changes across the parameter space, and a fixed choice leaves significant performance on the table.

This paper studies a single thesis:

\begin{quote}
\textbf{Checkpoint loading should be selected from workload shape.} The right backend depends on checkpoint layout, model scale, backend availability, and the layer at which loading is observed. A fixed rule such as ``always sync,'' ``always async,'' or ``always \texttt{io\_uring}'' is the wrong abstraction.
\end{quote}

The thesis is visible from two layouts. SafeTensors presents mostly whole-file or multi-shard eager loading: after metadata parsing, the loader must move large contiguous byte regions. ServerlessLLM-style partitioning presents a different workload: tensors are reconstructed from indexed partition files, creating many grouped range reads. These layouts are not cosmetic alternatives; they expose different I/O regimes. Scale then moves the boundary within a regime. A backend with low orchestration overhead can win for a small single-shard load, while a backend that sustains deeper submission queues can win for a large multi-shard load.

We built Tensora\footnote{Tensora is this paper's checkpoint-loading framework (Rust crate \texttt{tensora}, Python package \texttt{tensora\_py}). It is unrelated to Google's TensorStore library for array storage.}, an open framework for measuring these regimes. Tensora exposes synchronous POSIX I/O, Tokio-based async loading, Linux \texttt{io\_uring}, and a feature-driven \texttt{default} policy under shared format abstractions. We evaluate six public checkpoints from 0.6B to 32B parameters, two layouts, and three observation layers: Rust backend attribution, Python bindings, and a custom vLLM integration. The main environment is an H100 host with local NVMe and working \texttt{io\_uring}; H200 and A100 container environments with network-attached storage and blocked \texttt{io\_uring} test whether the policy degrades gracefully when the strongest H100 backend is unavailable.

The data support a regime claim, not a universal-loader claim. On H100 SafeTensors, explicit sync wins at the small end of the ladder, while explicit \texttt{io\_uring} wins from Qwen3-8B through Qwen3-32B. On H100 ServerlessLLM, sync is never fastest; async, \texttt{io\_uring}, and default read planning lead different rows. Python and vLLM reorder backend rankings because they measure different constructs, but the effect survives integration: in vLLM cold \texttt{load\_only} runs, Tensora's adaptive SafeTensors path reduces initialization by 37 to 48\% relative to vLLM's native loader for 8B and 32B checkpoints. Cross-environment results preserve the most stable structural fact (small SafeTensors loads favor sync) while showing that the large-scale winner depends on kernel and storage context.

\textbf{Contributions.}
\begin{enumerate}
  \item \textbf{Workload-shape abstraction.} We formalize checkpoint loading as a policy selection problem parameterized by format-level request geometry, model scale, and runtime capability, not peak device bandwidth alone. This abstraction identifies distinct performance regimes where different backends dominate, and regime boundaries that shift with checkpoint layout.
  \item \textbf{Capability-gated adaptive selector.} We design and evaluate a format-aware selector that uses measured workload features (total bytes, shard/partition count, average shard size, request structure) and runtime capability checks to choose among sync, async, and \texttt{io\_uring} backends. The selector degrades gracefully when backends are unavailable in restricted container environments.
  \item \textbf{Cross-layer evaluation.} We provide controlled evidence across three observation layers (Rust backend attribution, Python bindings, and a custom vLLM integration), showing that backend effects propagate through library and serving-engine boundaries. Integration layers reorder but do not erase backend differences.
  \item \textbf{Cross-environment validation.} Measurements across H100 (local NVMe, full backend set), H200 (network storage, \texttt{io\_uring} blocked), and A100 (network storage, \texttt{io\_uring} blocked) systems show that ordering regimes are more portable than backend identities.
  \item \textbf{Bounded operational policy.} We deliver a deliberately bounded recommendation: use a format-aware adaptive selector with explicit overrides and capability checks. The selector is not presented as an oracle; its misses are reported and used to identify where overrides and per-environment calibration are needed.
\end{enumerate}

\textbf{Roadmap.} Section~\ref{sec:related_work} positions the gap. Section~\ref{sec:design} describes Tensora and the adaptive policy. Section~\ref{sec:experimental_setup} defines methodology. Section~\ref{sec:results} presents results across three layers. Section~\ref{sec:discussion} interprets implications, Section~\ref{sec:limitations} bounds claims, and Section~\ref{sec:conclusion} concludes.
