\section{Results}
\label{sec:results}

The evaluation proceeds in three layers: Rust backend attribution (Section~\ref{sec:rust-results}), Python binding overhead (Section~\ref{sec:python-results}), and vLLM integration (Section~\ref{sec:vllm-results}). Tables present core evidence; figures visualize regime crossovers. Section~\ref{sec:oracle-analysis} compares the adaptive default against the best per-row explicit backend to bound the selector's effectiveness.

\subsection{Reading the Results}

The section follows the same control flow as the system. Rust results isolate backend attribution. Oracle analysis bounds the adaptive default against the best explicit backend on each row. Cross-environment results test capability-gated fallback. Python and vLLM results then show how the same choices behave once library and serving overhead are included.

\begin{table}[htbp]
\caption{Regime map: strongest observed policy by layout, environment, and scale. Color encodes the winning policy in the corresponding matrix. The key visual fact is non-dominance: every backend color appears, and the pattern changes with layout and environment.}
\label{tab:regime-map}
\centering
\small
\begin{tabular}{lllll}
\toprule
Layer / layout & 0.6B & 4B & 8B & 32B \\
\midrule
H100 SafeTensors & \syncwin & \syncwin & \uringwin & \uringwin \\
H100 ServerlessLLM & \uringwin & \asyncwin & \defaultwin & \defaultwin \\
H200 SafeTensors & \syncwin & \syncwin & \asyncwin & \asyncwin \\
H200 ServerlessLLM & \asyncwin & \asyncwin & \asyncwin & \defaultwin \\
A100 SafeTensors & \syncwin & \asyncwin & \asyncwin & \defaultwin \\
A100 ServerlessLLM & \defaultwin & \defaultwin & \defaultwin & \asyncwin \\
vLLM SafeTensors & \syncwin & \defaultwin & \defaultwin & \uringwin \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:regime-map} is the visual thesis of the paper. SafeTensors begins in the sync regime and shifts with scale. ServerlessLLM is not sync-led. Restricted environments replace \texttt{io\_uring} with async or default. vLLM preserves the small-versus-large structure while adding engine overhead. A single fixed backend cannot reproduce this map.

\begin{table}[htbp]
\caption{Effect-size summary for representative regime claims. Ratios are computed as slower time divided by faster time; larger values indicate stronger evidence for a regime distinction.}
\label{tab:effect-summary}
\centering
\small
\begin{tabular}{p{0.42\textwidth}p{0.24\textwidth}p{0.22\textwidth}}
\toprule
Claim & Evidence row & Effect size \\
\midrule
Small SafeTensors favors low-overhead sync/default over \texttt{io\_uring} & H100 Qwen3-0.6B & sync is 2.09$\times$ faster than \texttt{io\_uring}; default is 1.66$\times$ faster than sync \\
Large SafeTensors needs deeper submission & H100 Qwen3-32B & \texttt{io\_uring} is 1.28$\times$ faster than sync \\
Partitioning removes sync as winner & H100 ServerlessLLM Qwen3-32B & default is 1.44$\times$ faster than sync \\
\texttt{io\_uring} availability changes large-scale winner & H200 SafeTensors Qwen3-32B & async is 3.41$\times$ faster than sync when \texttt{io\_uring} is blocked \\
Serving impact survives integration & vLLM Qwen3-32B cold load\_only & Tensora default is 48.1\% faster than native \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Headline findings}

Four patterns emerge from the Rust and integration matrices. \textbf{First, SafeTensors exhibits a backend crossover.} Synchronous loading is the fastest explicit backend for model sizes up to 4B parameters; io\_uring is the fastest explicit backend from 8B parameters upward. No single fixed backend tracks both regimes.

\textbf{Second, the partitioned ServerlessLLM layout reverses the ranking.} Synchronous loading is never the fastest explicit backend for any model size in this ladder. The relevant trade-off is between async loading (strong on smaller and mid-sized range loads) and io\_uring (competitive at larger partition counts).

\textbf{Third, the adaptive default is useful but bounded.} It beats the best explicit backend in 3 of 12 H100 Rust cells (Table~\ref{tab:oracle}) and lands within roughly 3\% on several additional rows. The large misses are concentrated in medium-scale boundary regions. That pattern argues for adaptive defaults plus explicit overrides, not for hiding the selector or claiming oracle optimality.

\textbf{Fourth, integration overhead reorders but does not erase backend effects.} Python materialization makes absolute times much larger than Rust timings. vLLM adds engine initialization. Yet loader choice still changes cold initialization by tens of seconds at larger scales, which is the deployment-relevant outcome.

\textbf{Fifth, ordering regimes transfer better than backend names.} Cross-environment validation on H200 and A100 containers (where \texttt{io\_uring} is blocked by kernel security policy) confirms that small SafeTensors loads favor sync across all environments. The mid-to-large-scale winner changes with storage topology and kernel capability. This makes capability-gated adaptation more valuable, not less.

\subsection{Rust Backend Results}
\label{sec:rust-results}

The Rust harness is the authoritative backend-attribution layer because it measures checkpoint loading directly, without Python tensor conversion or serving-engine initialization costs. Unless stated otherwise, each matrix cell is one representative cold-cache run; Appendix~\ref{app:anchor-stats} reports five-repetition medians and min--max bands for selected anchors.

\subsubsection{SafeTensors Results}

Table~\ref{tab:rust-safetensors} summarises the explicit backend comparison for SafeTensors.

\begin{table}[htbp]
\caption{Rust cold-cache SafeTensors results (ms). The crossover from sync-favored small models to io\_uring-favored large multi-shard models is the evidential core behind the adaptive policy.}
\label{tab:rust-safetensors}
\centering
\begin{tabular}{lrrrr}
\toprule
Model & Sync & Async & Default & io\_uring \\
\midrule
Qwen3-0.6B & 555.28 & 1416.19 & 333.59 & 1159.24 \\
SmolLM3-3B & 1331.79 & 5802.05 & 3722.35 & 3812.35 \\
Qwen3-4B & 1820.85 & 5039.21 & 3083.18 & 2967.42 \\
Qwen3-8B & 3880.40 & 7074.40 & 3844.35 & 3172.37 \\
Qwen3-14B & 7030.25 & 13489.00 & 6138.70 & 5977.07 \\
Qwen3-32B & 15454.45 & 28763.46 & 12119.50 & 12089.05 \\
\bottomrule
\end{tabular}
\end{table}

Two regimes are visible. Sync dominates at the small end: 555~ms for Qwen3-0.6B versus 1159~ms for io\_uring. The ranking flips at Qwen3-8B: io\_uring at 3172~ms versus sync at 3880~ms, a 1.22$\times$ advantage. At the largest scale (Qwen3-32B), io\_uring is 1.28$\times$ faster than sync. The crossover is not a single threshold; it is a progression visible across the entire ladder.

Async is consistently the weakest backend for SafeTensors. Whole-file and multi-shard loads benefit from straightforward bulk parallelism rather than scheduler-mediated async orchestration, which adds overhead without corresponding I/O benefits on this workload shape.

\begin{figure}[htbp]
    \centering
    \begin{tikzpicture}
    \begin{axis}[
    width=0.88\textwidth,
    height=6cm,
    ybar,
    bar width=6pt,
    ymode=log,
    ylabel={Cold load time (ms)},
    xlabel={Model},
    symbolic x coords={0.6B,3B,4B,8B,14B,32B},
    xtick=data,
    legend style={at={(0.5,1.15)},anchor=south,legend columns=4},
    ymajorgrids=true,
    every axis plot/.style={fill=gray!40,draw=black},
    ]
    \addplot coordinates {(0.6B,555.28) (3B,1331.79) (4B,1820.85) (8B,3880.40) (14B,7030.25) (32B,15454.45)};
    \addplot coordinates {(0.6B,1416.19) (3B,5802.05) (4B,5039.21) (8B,7074.40) (14B,13489.00) (32B,28763.46)};
    \addplot coordinates {(0.6B,333.59) (3B,3722.35) (4B,3083.18) (8B,3844.35) (14B,6138.70) (32B,12119.50)};
    \addplot coordinates {(0.6B,1159.24) (3B,3812.35) (4B,2967.42) (8B,3172.37) (14B,5977.07) (32B,12089.05)};
    \legend{sync,async,default,io\_uring}
    \end{axis}
    \end{tikzpicture}
    \caption{SafeTensors backend crossover (log scale). Sync dominates the left (small models); io\_uring dominates the right (large multi-shard models). Default trails on medium-scale rows where the heuristic's feature mapping is imprecise.}
    \label{fig:safetensors-crossover}
\end{figure}

Figure~\ref{fig:safetensors-crossover} visualizes the progression. \texttt{default} adds planning (chunking, coalescing) that helps on Qwen3-0.6B (333~ms, beating all explicit backends) but misses the medium-scale crossover region on SmolLM3-3B and Qwen3-4B.

\subsubsection{ServerlessLLM Results}

The ServerlessLLM layout produces a different ordering, shown in Table~\ref{tab:rust-serverlessllm}.

\begin{table}[htbp]
\caption{Rust cold-cache ServerlessLLM results (ms). Synchronous loading never wins; the ranking shifts from async at smaller partition scales toward io\_uring at larger sizes.}
\label{tab:rust-serverlessllm}
\centering
\begin{tabular}{lrrrr}
\toprule
Model & Sync & Async & Default & io\_uring \\
\midrule
Qwen3-0.6B & 1645.06 & 1471.34 & 1432.25 & 1387.57 \\
SmolLM3-3B & 3941.88 & 2391.24 & 3182.28 & 3091.49 \\
Qwen3-4B & 5137.72 & 3504.25 & 3551.82 & 3727.14 \\
Qwen3-8B & 11048.40 & 9463.37 & 7589.47 & 9043.51 \\
Qwen3-14B & 18661.25 & 16139.08 & 13422.94 & 13257.23 \\
Qwen3-32B & 39822.09 & 29180.89 & 27659.78 & 29779.86 \\
\bottomrule
\end{tabular}
\end{table}

Sync is the slowest explicit backend on every ServerlessLLM row, between 1.1$\times$ (Qwen3-0.6B) and 1.4$\times$ (Qwen3-32B) behind the strongest alternative. The best explicit backend varies with scale: async dominates at smaller partition counts (SmolLM3-3B, Qwen3-4B), while io\_uring becomes competitive at larger scales (Qwen3-14B, Qwen3-32B).

\texttt{default} outperforms all explicit backends on Qwen3-8B (7589~ms vs 9044~ms for the next best, io\_uring) and Qwen3-32B (27660~ms vs 29181~ms for async) because its read planning---coalescing nearby ranges, dynamic chunk sizing---benefits partitioned workloads more than any single explicit submission policy. It trails explicit async on SmolLM3-3B (3182~ms vs 2391~ms, a 1.33$\times$ gap), the same medium-scale fragility seen in SafeTensors.

\begin{figure}[htbp]
    \centering
    \begin{tikzpicture}
    \begin{axis}[
    width=0.88\textwidth,
    height=6cm,
    ybar,
    bar width=6pt,
    ymode=log,
    ylabel={Cold load time (ms)},
    xlabel={Model},
    symbolic x coords={0.6B,3B,4B,8B,14B,32B},
    xtick=data,
    legend style={at={(0.5,1.15)},anchor=south,legend columns=4},
    ymajorgrids=true,
    every axis plot/.style={fill=gray!40,draw=black},
    ]
    \addplot coordinates {(0.6B,1645.06) (3B,3941.88) (4B,5137.72) (8B,11048.40) (14B,18661.25) (32B,39822.09)};
    \addplot coordinates {(0.6B,1471.34) (3B,2391.24) (4B,3504.25) (8B,9463.37) (14B,16139.08) (32B,29180.89)};
    \addplot coordinates {(0.6B,1432.25) (3B,3182.28) (4B,3551.82) (8B,7589.47) (14B,13422.94) (32B,27659.78)};
    \addplot coordinates {(0.6B,1387.57) (3B,3091.49) (4B,3727.14) (8B,9043.51) (14B,13257.23) (32B,29779.86)};
    \legend{sync,async,default,io\_uring}
    \end{axis}
    \end{tikzpicture}
    \caption{ServerlessLLM backend ordering (log scale). Sync is never the winner; the layout shifts the ranking toward async and io\_uring. Default benefits from read planning on large partitioned loads.}
    \label{fig:serverlessllm-crossover}
\end{figure}

Figure~\ref{fig:serverlessllm-crossover} shows the ranking shift. The figure confirms that the ServerlessLLM workload shape---many small range reads across partition files---rewards different backend policies than the SafeTensors contiguous regime.

\subsection{Oracle analysis: how well does the adaptive default track the best backend?}
\label{sec:oracle-analysis}

The adaptive default is not expected to match the best per-row explicit backend in every cell---it is a heuristic, not an oracle. Table~\ref{tab:oracle} quantifies the gap.

\begin{table}[htbp]
\caption{Default versus per-row best explicit backend. Cells show \texttt{default} / \texttt{best\_explicit} ratio. Ratios $<1.0$ mean default wins; ratios $>1.0$ mean default trails.}
\label{tab:oracle}
\centering
\begin{tabular}{lrr}
\toprule
Model & SafeTensors & ServerlessLLM \\
\midrule
Qwen3-0.6B & 0.60 & 1.03 \\
SmolLM3-3B & 2.79 & 1.33 \\
Qwen3-4B & 1.69 & 1.01 \\
Qwen3-8B & 1.21 & 0.80 \\
Qwen3-14B & 1.03 & 1.01 \\
Qwen3-32B & 1.00 & 0.93 \\
\bottomrule
\end{tabular}
\end{table}

Default beats the best explicit backend in 3 of 12 cells and lands within roughly 3\% on four additional cells. For SafeTensors, default correctly routes the smallest model (Qwen3-0.6B) to a fast path and tracks the large-scale io\_uring regime within 3\% at Qwen3-14B and near parity at Qwen3-32B. The medium-scale misses---SmolLM3-3B (2.79$\times$), Qwen3-4B (1.69$\times$), and Qwen3-8B (1.21$\times$)---are the region where the crossover from sync to io\_uring is most sensitive to per-checkpoint shard structure that the current feature set does not fully resolve.

For ServerlessLLM, default beats the best explicit backend in 2 of 6 cells and is close on most remaining rows. The single notable miss is SmolLM3-3B (1.33$\times$), where async's grouped range execution fits the partition structure better than default's chosen path. The remaining ratios are near-parity ($\leq 1.03$).

\textbf{Interpretation.} The misses are not hidden; they are the point at which the adaptive policy becomes an engineering policy rather than an oracle. They identify where the feature mapping is fragile and where explicit overrides (for example, force sync for SafeTensors SmolLM3-3B; force async for ServerlessLLM SmolLM3-3B) provide clear value. A deployable selector should therefore expose overrides instead of pretending every boundary can be learned from one static threshold.

\subsection{Cross-Environment Validation}
\label{sec:cross-env}

The H100 results above were collected with all four backends available, including \texttt{io\_uring}. Two additional environments---an H200 container and an A100 container, both with network-attached storage---block \texttt{io\_uring} at the kernel level (\texttt{EPERM}). This constraint tests two questions: (1) whether the adaptive selector degrades gracefully when its strongest large-scale backend is unavailable, and (2) whether backend ordering regimes transfer across hardware.

\subsubsection{H200 Results (io\_uring blocked)}

Table~\ref{tab:h200-safetensors} shows the H200 SafeTensors matrix. With \texttt{io\_uring} blocked, async replaces it as the fastest backend at large scale, while sync remains dominant for small models.

\begin{table}[htbp]
\caption{H200 cold-cache SafeTensors results (ms). \texttt{io\_uring} is blocked (\texttt{EPERM}). Sync dominates at small scale; async takes over at large scale in the absence of \texttt{io\_uring}.}
\label{tab:h200-safetensors}
\centering
\begin{tabular}{lrrr}
\toprule
Model & Sync & Async & Default \\
\midrule
Qwen3-0.6B & 604.34 & 852.17 & 649.23 \\
SmolLM3-3B & 2013.66 & 2938.30 & 2130.51 \\
Qwen3-4B & 909.83 & 2478.10 & 941.24 \\
Qwen3-8B & 3924.69 & 2977.59 & 3213.01 \\
Qwen3-14B & 25872.03 & 4767.67 & 4906.13 \\
Qwen3-32B & 36205.91 & 10623.83 & 12356.23 \\
\bottomrule
\end{tabular}
\end{table}

The H200 SafeTensors crossover preserves the structural pattern from H100: sync wins at small scale (0.6B--4B), while a different backend dominates at large scale (8B+). On H100 that backend is \texttt{io\_uring}; on H200 it is async. The selector adapts: \texttt{default} tracks within 3--8\% of the best available backend for 0.6B--14B, though it trails async by 16\% at 32B.

Table~\ref{tab:h200-serverlessllm} shows the H200 ServerlessLLM matrix.

\begin{table}[htbp]
\caption{H200 cold-cache ServerlessLLM results (ms). Async dominates most rows; \texttt{default} beats both explicit backends at 32B.}
\label{tab:h200-serverlessllm}
\centering
\begin{tabular}{lrrr}
\toprule
Model & Sync & Async & Default \\
\midrule
Qwen3-0.6B & 1565.27 & 1491.44 & 1586.84 \\
SmolLM3-3B & 3945.68 & 1947.39 & 1987.07 \\
Qwen3-4B & 5100.25 & 2489.27 & 2809.79 \\
Qwen3-8B & 10788.88 & 6045.47 & 6098.06 \\
Qwen3-14B & 17765.28 & 14450.40 & 15453.47 \\
Qwen3-32B & 39061.57 & 27367.74 & 24259.15 \\
\bottomrule
\end{tabular}
\end{table}

The ServerlessLLM ordering on H200 matches H100 structurally: sync is never the fastest, and async dominates most rows. \texttt{default} outperforms both explicit backends at 32B (24259~ms vs 27368~ms for async), consistent with the H100 finding that read planning benefits partitioned loads at large scale.

\subsubsection{A100 Results (io\_uring blocked, reduced ladder)}

The A100 environment runs a four-model reduced ladder (0.6B, 4B, 8B, 32B) with the same \texttt{EPERM} constraint. Table~\ref{tab:a100-safetensors} and Table~\ref{tab:a100-serverlessllm} summarise the results.

\begin{table}[htbp]
\caption{A100 cold-cache SafeTensors results (ms). Network-attached storage with AMD EPYC CPU. Sync wins at 0.6B; async dominates at 4B--8B; sync recovers at 32B.}
\label{tab:a100-safetensors}
\centering
\begin{tabular}{lrrr}
\toprule
Model & Sync & Async & Default \\
\midrule
Qwen3-0.6B & 724.26 & 1094.43 & 965.95 \\
Qwen3-4B & 7598.41 & 3532.11 & 7810.53 \\
Qwen3-8B & 8412.36 & 4235.21 & 11896.57 \\
Qwen3-32B & 46923.42 & 60807.34 & 43816.68 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[htbp]
\caption{A100 cold-cache ServerlessLLM results (ms). \texttt{default} wins at 0.6B, 4B, and 8B; async dominates at 32B.}
\label{tab:a100-serverlessllm}
\centering
\begin{tabular}{lrrr}
\toprule
Model & Sync & Async & Default \\
\midrule
Qwen3-0.6B & 2183.33 & 2250.91 & 1554.97 \\
Qwen3-4B & 7551.79 & 4486.75 & 3523.25 \\
Qwen3-8B & 14707.01 & 11528.26 & 11431.61 \\
Qwen3-32B & 56025.98 & 40472.54 & 44854.60 \\
\bottomrule
\end{tabular}
\end{table}

The A100 SafeTensors results show a distinctive pattern: sync wins at 0.6B, async dominates the mid-range (4B--8B), but sync recovers at 32B. This contrasts with both H100 (where \texttt{io\_uring} dominates large scale) and H200 (where async dominates large scale). The \texttt{default} selector on A100 performs well at 0.6B and 32B but misses the mid-range async regime, producing its largest gap at 8B (2.81$\times$ behind async). For ServerlessLLM, \texttt{default} wins at 0.6B and 4B and tracks within 1\% of async at 8B.

\subsubsection{Cross-Environment Summary}

Table~\ref{tab:cross-env-best} summarises the best explicit backend per row across all three environments.

\begin{table}[htbp]
\caption{Best explicit backend by environment for SafeTensors. The crossover from sync to another backend persists on H100 and H200; A100 shows a non-monotonic pattern where sync recovers at 32B. Which alternative wins at mid-to-large scale depends on hardware and storage topology.}
\label{tab:cross-env-best}
\centering
\begin{tabular}{llll}
\toprule
Model & H100 & H200 & A100 \\
\midrule
Qwen3-0.6B & sync & sync & sync \\
Qwen3-4B & sync & sync & async \\
Qwen3-8B & io\_uring & async & async \\
Qwen3-32B & io\_uring & async & sync \\
\bottomrule
\end{tabular}
\end{table}

Three findings from the cross-environment comparison: (1)~Sync dominates small SafeTensors models in all three environments---this is the most portable ordering. (2)~At large scale, the strongest backend depends on available kernel interfaces and storage topology: \texttt{io\_uring} on local NVMe (H100), async on network storage (H200), and sync on a different network-storage configuration (A100 at 32B). (3)~The adaptive selector degrades gracefully: when \texttt{io\_uring} is unavailable, it produces within-regime-competitive results on most rows, confirming that the capability-gated fallback mechanism works as designed.

\subsection{Python Results}
\label{sec:python-results}

Python benchmarks were collected on Qwen/Qwen3-8B to quantify user-facing overhead and to evaluate whether binding optimisation improved practical performance.

The most important Python result is that binding engineering matters. Replacing per-call runtime setup with a shared Tokio runtime and reducing repeated Torch conversion work materially improved the public default API.

\begin{table}[htbp]
\caption{Python eager-load times for Qwen3-8B (seconds). Binding tuning and selector redesign both contributed measurable improvements.}
\label{tab:python-optimisation}
\centering
\begin{tabular}{lrrr}
\toprule
Case & Baseline bindings & After binding tuning & After selector tuning \\
\midrule
SafeTensors default & 15.88 & 14.55 & 11.55 \\
ServerlessLLM default & 22.45 & 20.50 & 18.45 \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:python-optimisation} shows the sequence. For SafeTensors on Qwen3-8B, the default path moves from 15.88~s to 11.55~s, a 27.3\% reduction from the baseline. For ServerlessLLM, the same sequence moves default from 22.45~s to 18.45~s, a 17.8\% reduction.

The Python numbers are much larger than the Rust-only numbers (11.55~s vs 3.84~s for SafeTensors default) because eager loading at the binding layer includes substantial tensor materialization and Python object construction. The key conclusion: Python is not a transparent pass-through layer. Backend improvements matter, but they are filtered through the cost of converting checkpoint contents into Torch tensors and Python dictionaries.

\subsection{vLLM Results}
\label{sec:vllm-results}

The final vLLM matrix completed successfully across all six models, all benchmark loaders, warm and cold cache conditions, and all three benchmark kinds.

\textbf{Stock vLLM versus Tensora.} The \texttt{native} loader follows vLLM's default weight-loading path; \texttt{ts\_*} loaders use Tensora-backed integrations. On cold \texttt{load\_only}, Tensora's default SafeTensors path reduces initialization time relative to \texttt{native} across all measured scales (Table~\ref{tab:vllm-baseline}).

\begin{table}[htbp]
\caption{Cold vLLM load\_only: native vs. Tensora (ms). The gap is an end-to-end effect (CUDA context, memory pools, and loader policy), not a pure Rust microbenchmark.}
\label{tab:vllm-baseline}
\centering
\begin{tabular}{lrrr}
\toprule
Model & \texttt{native} & \texttt{ts\_safetensors\_default} & $\Delta$ \\
\midrule
Qwen3-0.6B & 23659 & 20578 & $-13.0\%$ \\
Qwen3-8B & 56639 & 35556 & $-37.2\%$ \\
Qwen3-32B & 169971 & 88200 & $-48.1\%$ \\
\bottomrule
\end{tabular}
\end{table}

The improvement grows with model size: 13\% for the smallest model, 37\% at 8B, and 48\% at 32B. This is consistent with the Rust SafeTensors finding (io\_uring advantage grows at larger scales) filtered through vLLM's fixed initialization overhead.

For cold load\_only across all models, the best loader varies with model size. Table~\ref{tab:vllm-load-only} shows the strongest loader in the final cold-load-only configuration.

\begin{table}[htbp]
\caption{Best cold load\_only vLLM loader by model. The pattern follows the Rust SafeTensors regime: sync-style loading at small scales, io\_uring at large scales.}
\label{tab:vllm-load-only}
\centering
\begin{tabular}{lrl}
\toprule
Model & Best init time (ms) & Best loader \\
\midrule
Qwen3-0.6B & 20417 & ts\_safetensors\_sync \\
SmolLM3-3B & 27334 & ts\_safetensors\_default \\
Qwen3-4B & 28475 & ts\_safetensors\_default \\
Qwen3-8B & 35556 & ts\_safetensors\_default \\
Qwen3-14B & 47751 & ts\_safetensors\_io\_uring \\
Qwen3-32B & 88164 & ts\_safetensors\_io\_uring \\
\bottomrule
\end{tabular}
\end{table}

The vLLM best-loader pattern mirrors the Rust SafeTensors crossover: smaller models favor sync-style loaders, larger models favor default or io\_uring. The match is not exact---vLLM initialization includes fixed costs that dilute the pure-I/O ranking---but the regime structure survives.

For TTFT (Table~\ref{tab:vllm-ttft}), the ranking is broadly similar to load\_only but does not match row-for-row because TTFT includes the first generation step after initialization.

\begin{table}[htbp]
\caption{Best cold TTFT vLLM loader by model.}
\label{tab:vllm-ttft}
\centering
\begin{tabular}{lrl}
\toprule
Model & Best TTFT (ms) & Best loader \\
\midrule
Qwen3-0.6B & 20273 & ts\_safetensors\_sync \\
SmolLM3-3B & 27463 & ts\_safetensors\_sync \\
Qwen3-4B & 28530 & ts\_safetensors\_default \\
Qwen3-8B & 35100 & ts\_safetensors\_default \\
Qwen3-14B & 48202 & ts\_safetensors\_default \\
Qwen3-32B & 88551 & ts\_safetensors\_default \\
\bottomrule
\end{tabular}
\end{table}

The SafeTensors loader family consistently outperforms the ServerlessLLM family for both metrics. Load\_only and TTFT show backend effects mainly in initialization; decode is comparatively insensitive once weights are resident.

\subsection{Layer summary}

Four principles emerge. (1) Backend ranking depends on workload shape: SafeTensors favors sync at small scale and \texttt{io\_uring} at large H100 scale, while ServerlessLLM favors async, \texttt{io\_uring}, or default planning and disfavors sync. (2) The adaptive default is useful but bounded: it wins or ties several stable-regime rows and misses medium-scale rows that need explicit overrides. (3) Integration layers reorder but do not eliminate backend differences: Python adds tensor conversion overhead and vLLM adds initialization cost, yet loader choice still changes user-visible startup. (4) Cross-environment validation confirms that sync dominates small SafeTensors models everywhere; the mid-to-large-scale winner varies with kernel interface availability and storage topology, and A100 shows a non-monotonic pattern where sync/default recovers at the largest scale. The selector's capability-gated fallback adapts to these constraints better than a hard-coded \texttt{io\_uring}-first policy would.